### 2019-07-09 version_2/final_response_version_1.tex
Wrote intro to final response, revised summary and conclusion. ToDo - check if all issues have been addressed. Revise response RC2.

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
%\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{url}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\thispagestyle{empty}

\begin{document}
\section*{Final response}
We thank both referees for their thorough review of our paper and the evaluation of the formulation of the dry deposition scheme.
In consideration of all comments, it became necessary to revise a major part of the dry deposition implementation in our model. Particularly, we  
\begin{itemize}
  \item have addressed and corrected the used equations and the implementation of the mosaic approach in the model,
  \item have fixed all related issues in the manuscript,
  \item have included a diagnostic output of dry deposition velocities,
  \item now refer to our new dry deposition scheme as \emph{mOSaic scheme},
  \item have refined the definition of our model experiments to make them more consistent, especially
    \begin{itemize}
    \item using SWVL3 and
    \item CEDS emissions in accordance to the meteorological year.
    \end{itemize}
    This also includes replacement or removal of some previous experiments
    \begin{itemize}
    \item \emph{EMEP\_SWVL4} $\rightarrow$ \emph{mOSaic\_SWVL1},
    \item \sout{\emph{EMEP\_ppgs}},
    \item \sout{\emph{EMEP\_ppgssh}}, and
    \item \emph{EMEP\_ppgs\_2005} $\rightarrow$ \emph{mOSaic\_emis2014}.
    \end{itemize}
    We have added two experiments in which we explicitly change the prescribed ozone surface resistance $R^\mathrm{O_3}$:
    \begin{itemize}
    \item \emph{mOSaic\_desert},
    \item\emph{mOSaic\_hough}.
    \end{itemize}
  \item have elaborated on the evaluation and discussion of the results and included a section in which we compare our results to the MACC-reanalysis.
\end{itemize}
In the following, we include a detailed account of all referee comments and the track-change-version of the manuscript. We have not updated the responses already published in gmdd (RC1, RC3). Hence, the referenced numbers will be different in the final manuscript.
\newpage
\input{response_version_1}
\input{response_version_1_rc2}
\input{response_version_1_rc3}
\newpage

\end{document}