### 2019-07-11 version_2/final_response_version_1.tex
Small adjustments to response. Added Diff.

% Authors' response
%\documentclass{scrartcl}
%\usepackage[utf8]{inputenc}
%\usepackage[english]{babel}
%\usepackage[T1]{fontenc}
%\usepackage{amsmath}
%\usepackage{xcolor}
%\usepackage{url}
%\usepackage{multirow}
%\usepackage{graphicx}
%\usepackage{hyperref}
%\usepackage[normalem]{ulem}
%\thispagestyle{empty}

%\begin{document}
\section*{Final response}
We thank both referees for their thorough review of our paper and the evaluation of the formulation of the dry deposition scheme.
In consideration of all comments, it became necessary to revise a major part of the dry deposition implementation in our model. Particularly, we  
\begin{itemize}
\item have addressed and corrected the equations in
  \begin{itemize}
  \item the implementation of the mosaic approach,
  \item the computation of $R_a$.
  \end{itemize}
  \item have fixed the equations with respect to comprehensibility (e.g., indexing),
  \item have included a diagnostic output of dry deposition velocities,
  \item now refer to our new dry deposition scheme as \emph{mOSaic scheme},
  \item have refined the definition of our model experiments to make them more consistent, especially
    \begin{itemize}
    \item using the soil moisture at a more appropriate level (SWVL3) and
    \item CEDS emissions in accordance to the meteorological year.
    \end{itemize}
    This also includes replacement or removal of some previous experiments
    \begin{itemize}
    \item \emph{EMEP\_SWVL4} $\rightarrow$ \emph{mOSaic\_SWVL1},
    \item \sout{\emph{EMEP\_ppgs}},
    \item \sout{\emph{EMEP\_ppgssh}}, and
    \item \emph{EMEP\_ppgs\_2005} $\rightarrow$ \emph{mOSaic\_emis2014}.
    \end{itemize}
    We have added two experiments in which we explicitly change the prescribed ozone surface resistance $R^\mathrm{O_3}$:
    \begin{itemize}
    \item \emph{mOSaic\_desert},
    \item\emph{mOSaic\_hough}.
    \end{itemize}
  \item have elaborated on the evaluation and discussion of the results and
  \item included a section in which we compare our results to the MACC-reanalysis.
\end{itemize}
In the following, we account for all referee comments in detail and include the track-change-version of the manuscript. {\bf Remark:} We have not updated the responses that have been already published in gmdd (RC1, RC3). Hence, the referenced values therein differ from the final manuscript.
\newpage
\input{version_2/response_version_1}
\input{version_2/response_version_1_rc2}
\input{version_2/response_version_1_rc3}
\newpage

%\end{document}